Miniproject - Combinatorics

### 2021-12-20 Combinatorics/submission.tex
Use candidate number instead of student number
Finalized combinatorics

\title{Miniproject - Combinatorics}
\author{Simon Pohmann}
\date{}

@@

    Consider any $W = \{ (1, i_1, a_1), ..., (1, i_w, a_w) \} \subseteq V_1$ and let $F = \{ a_1, ..., a_w \}$.
    Then
    \begin{equation*}
        W \subseteq [t] \times F \ \Rightarrow \ |W| \leq t|F|
    \end{equation*}
    By choice of $C_1, ..., C_t$, we have that the number of $j$ with $x \in C_j$ is the same for all $x \in L_{n - 1}$, say $k$.
    Since the $C_i$ are maximal chains, each intersects $L_{n - 1}$ in exactly one element.
    So have bijection
    \begin{align*}
        \{ (x, i) \ | \ x \in L_{n - 1}, 1 \leq i \leq t, x \in C_i\} \to \{ i \ | \ 1 \leq i \leq t \}, \quad (x, i) \mapsto i
    \end{align*}
    where the set on the left-hand side has size $k |L_{n - 1}|$. It follows that $k|L_{n - 1}| = t$.

    Since $\max C_j \in L_{n - 1}$ for all $j$, we have
    \begin{align*}
        N(W) =& \{ (2, j, b) \ | \ b \in L_n, \exists a \in F: \max C_j < a \} \\
        =& \{2\} \times \{ j \ | \ \exists a \in F: \max C_j < a \} \times L_n \\
        =& \{2\} \times \{ j \ | \ \max C_j \in \partial F \} \times L_n
    \end{align*}
    and so by the above
    \begin{align*}
        |N(W)| = |L_n| \sum_{\max C_j \in \partial F} 1 = |L_n| \sum_{x \in \partial F} \sum_{x \in C_j} 1 = |L_n| \sum_{x \in \partial F} k = |L_n| |\partial F| k
    \end{align*}
    Together with the assumption that $|\partial F| |L_n| \geq |F| |L_{n - 1}|$ we see that
    \begin{equation*}
        |W| \leq t |F| = k |L_{n - 1}| |F| \leq k |\partial F| |L_n| = |N(W)|
    \end{equation*}
    So Hall's theorem yields a matching $M \subseteq E$ from $V_1$ to $V_2$.
    As $|V_1| = |V_2|$, this is already a 1-to-one correspondence.

@@

    Note that this is the partition into \emph{exactly} $k$ sets; 
    I am not completely sure what the exercise description tells us to show, but if it is about the number of ways of decomposing $[m]$ into \emph{at most} $k$ sets, then have
    \begin{equation*}
        \Sigma S(m, k)(2) = 2^{m - 1} \quad \text{and} \quad \Sigma S(m, k)(3) = 3^{m - 1} = \frac 1 2 3^{m - 1} + \frac 1 2
    \end{equation*}
    for $m \geq 3$.
    Here $\Sigma S(m, k) = \sum_{l \leq k} S(m, l)$ is the number of partitions of $[m]$ into at most $k$ sets.
    \begin{proof}
        By Prop.~\ref{prop:partitions_recursion} have that for $m \geq 1$ we have $S(m, 1) = 1$.
        Further, for $m \geq 2$ have
        \begin{equation*}
            S(m, 2) = \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 = 2^{m - 1} - 1
        \end{equation*}
        Applying this once more yields for $m \geq 3$ that
        \begin{align*}
            S(m, 3) =& \sum_{n = 1}^{m - 1} {m - 1 \choose n} (2^{n - 1} - 1) = \frac 1 2 \sum_{n = 1}^{m - 1} {m - 1 \choose n} 2^n - \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 \\
            =&\frac 1 2 \sum_{n = 0}^{m - 1} {m - 1 \choose n} 2^n - \frac 1 2 - \sum_{n = 0}^{m - 1} {m - 1 \choose n} 1 + 1 \\
            =& \frac 1 2 (1 + 2)^{m - 1} - 2^{m - 1} + \frac 1 2 = \frac 1 2 (3^{m - 1} - 2^m + 1)
        \end{align*}
    \end{proof}
    Let $m$ be even and consider $\mathcal{A} \subseteq \Pi_m$ of partitions of $[m]$ into two equally sized sets.
    Then
    \begin{equation*}
        |\mathcal{A}| = \frac 1 2 {m \choose m/2}
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{equation*}
            f: [m]^{(m/2)} \to \mathcal{A}, \quad A \mapsto \{ A, [m] \setminus A \}
        \end{equation*}
        Then for $A$ and $B \neq A, [m] \setminus A$ have that $\{ A, [m] \setminus A \} \neq \{ B, [m] \setminus B \}$.
        Conversely, $A$ and $[m] \setminus A$ have the same image under $f$.
        This shows that $f$ is 2-to-1 and thus 
        \begin{equation*}
            |\mathcal{A}| = \frac 1 2 \left| [m]^{(m/2)} \right| = \frac 1 2 {m \choose m/2}
        \end{equation*}
    \end{proof}
    Further have
    \begin{equation*}
        |\partial \mathcal{A}| = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{align*}
            g: \underbrace{\{ (A, B, C) \ | \ A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m \}}_{ =: \mathcal{G} } \to \partial \mathcal{A}, \quad (A, B, C) &\mapsto \{ A, B, C \}
        \end{align*}
        Note that $g$ is well-defined, as for $A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m$ have that $\pi := \{ A, B, C \}$ is a partition in $L_{m - 3}$ that refines the partition $\{ A, B \cup C \} \in \mathcal{A} \subseteq L_{m - 2}$.
        Further, it is easy to see that $g$ is surjective, as every partition $\pi \in \partial \mathcal{A}$ satisfies $|\pi| = 3$ and has some $A \in \pi$ with $|A| = m/2$.

@@

\section{Part VI}
First of all, in Example~\ref{ex:partitions_have_no_lym} we have already shown that the poset $\Pi_m$ does not satisfy the conditions from Prop.~\ref{prop:characterization_lym}.
Now we want to investigate whether $\mathcal{P}_{k, d}$ does. 
\begin{lemma}
    \label{prop:lp_algorithm}
    Let $P$ be a poset with layers $L_0, ..., L_d$.
    Fix $0 < l \leq d$ and let $L_{l - 1} = \{ u_1, ..., u_m \}, L_l = \{ v_1, ..., v_n \}$.

    Assume there exists a matrix $A = (a_{ij}) \in \Q_{\geq 0}^{m \times n}$ such that
    \footnote{We use $1 = (1, ..., 1)^T$ as the vector consisting of $1$ only here.}
    \begin{equation*}
        1^T A 1 = 1
    \end{equation*}
    and for all $i, j$
    \begin{equation*}
        e_i^T A 1 = \frac 1 {|L_0|} \quad \text{and} \quad 1^T A e_j = \frac 1 {|L_1|}
    \end{equation*}
    and $a_{ij} = 0$ whenever $u_i \not< v_j$ holds.

@@

    Now consider a (finite) sequence of maximal chains $(C_k)_{k \in I}$ such that the chain corresponding to $u_i < v_j$ occurs $a_{ij} |I|$ times.
    By the assumption, we know that for each $i$ have
    \begin{equation*}
        |\{k \ | \ u_i \in C_k\}| = \sum_{u_i < v_j} a_{ij}|I| = \frac 1 {|L_0|} |I|
    \end{equation*}
    and so each element $u_i \in L_{l - 1}$ is contained in the same number $|I| / |L_0|$ of chains $C_k$.

    Similarly, for each $J$ have
    \begin{equation*}
        |\{k \ | \ v_j \in C_k\}| = \sum_{v_j > u_i} a_{ij}|I| = \frac 1 {|L_1|} |I|
    \end{equation*}
    The claim follows.
\end{proof}
This lemma does not seem super useful at first, but it gives us a way to check in time polynomial in $O(|P|)$ whether a given poset $P$ satisfies the conditions from Prop.~\ref{prop:characterization_lym}, by using an LP solver.
Note that this is not easily possible when working directly with any of the three conditions defined before.

@@

    If there exists a sequence of maximal chains $C'_1, ..., C'_t$ in $\mathcal{C}(P)$ such that for a chain $C$ chosen uniformly at random from among them, have
    \begin{equation*}
        \Pr[A \in C] = \frac {|A|} {|L_{r(A)}|}
    \end{equation*}
    for all $A \in \mathcal{C}(P)$, then $P$ satisfies the conditions of Prop.~\ref{prop:characterization_lym}.
\end{proposition}
\begin{proof}
    Exactly as in Prop.~\ref{prop:core_of_multisets}.
\end{proof}
I still do not know how to use this to prove that $\mathcal{P}_{k, d}$ does indeed satisfy Prop.~\ref{prop:characterization_lym}, but at least it provides a way for an even faster algorithm to test it for specific $k, d$.
The idea is just that we can adjust the LP algorithm from Lemma~\ref{prop:lp_algorithm} to check the scaled variant, and so we get a polynomial time algorithm in $|\mathcal{C}_{k, d}|$.
In particular, we have the following.
\begin{lemma}
    Let $L_{l - 1} = \{u_1, ..., u_m\}$ and $L_l = \{v_1, ..., v_n\}$ be two consecutive layers of $\mathcal{C}_{k, d}$.
    Assume there exists a matrix $A = (a_{ij}) \in \Q_{\geq 0}^{m \times n}$ such that 
    \begin{equation*}
        1^T A 1 = 1
    \end{equation*}
    and for all $i, j$
    \begin{equation*}
        e_i^T A 1 = \frac {|P(u_i)|} {\sum_{u \in L_{l - 1}} |P(u)|} \quad \text{and} \quad 1^T A e_j = \frac {|P(v_j)|} {\sum_{v \in L_l} |P(v)|}
    \end{equation*}
    and $a_{ij} = 0$ whenever $u_i \not< v_j$ holds.

@@

    \paragraph{Case 3} If $M = (k_1, ..., k_d)$, see that $2l - 1 = \sum_i k_i$ and so the layers $L_{l - 1}, L_l$ must be the two unique middle layers.
    Now note that $|L_l| = |L_{l - 1}|$ and so it suffices to show that $|\partial F| \geq |F|$.
    However, this follows directly by the symmetric chain decomposition of the poset $P$.
    We have proven the symmetric chain decomposition only for $k_1 = ... = k_d$, but the general version can be found in \cite{anderson}.
\end{proof}
There are two parts I find interesting in this proof.
\begin{itemize}
    \item First of all, the idea to consider all $\bigtimes_i \{ 0, ..., k_i \}$ with possibly distinct $k_i$ makes it easy to use induction, as it suffices to decrease just one coordinate to use the induction hypothesis.
    \item The decomposition
    \begin{equation*}
        L_{i - 1} \times F = \bigcup_{M \in P', r(M) = 2i - 1} \{ (u, a) \in L_{i - 1} \times F \ | \ u + a = M \}
    \end{equation*}
    is just very clever, as it really transforms the set of tuples $(u, a)$ into multiple sets that are more or less $\{ a \in F \ | \ a \leq M \}$ and much easier to handle.
\end{itemize}
Interestingly, I have also found a variant of the LYM inequality (``pseudo-LYM inequality'') for the poset $\Pi_m$ that we studied earlier.
\begin{proposition}
    \label{prop:lym_variant_partitions}
    Let $\mathcal{A} \subseteq \Pi_m$ be a nonempty antichain. Then define
    \begin{equation*}
        a(n_1, ..., n_k) := |\bigl\{ A \in \mathcal{A} \ \bigm| \ |A| = k, \ |\{ S \in A \ | \ |S| = j \}| = n_j \bigr\}|
    \end{equation*}
    the number of partitions $A \in \mathcal{A}$ such that the classes in $A$ have sizes $n_1, ..., n_k$.
    Then
    \begin{equation*}
        \sum_{k \geq 1} \sum_{n_1, ..., n_k} \frac {a(n_1, ..., n_k) k!} {{m \choose n_1, ..., n_k} {m - 1 \choose k - 1}} \leq 1
    \end{equation*}
\end{proposition}
\begin{proof}
    See \cite{lym_partitions}.
\end{proof}
Note that the total number of partitions $\pi$ such that the classes in $\pi$ have sizes $n_1, ..., n_k$ is given by an expression involving the multinomial coefficient, namely
\begin{equation*}
    |\bigl\{ \pi \in \Pi_m \ \bigm| \ |\pi| = k, |\{ S \in \pi \ | \ |S| = j\}| = n_j \bigr\}| = {m \choose n_1, ..., n_k} \underbrace{ \prod_i \frac 1 {|\{ j \ | \ n_j = i\}|!} }_{=: \gamma(n_1, ..., n_k)}
\end{equation*}
Intuitively, the reason is that there are $m!$ ways to order the elements $[m]$, and after splitting such a sequence in chunks of sizes $n_1, ..., n_k$, there are $n_1! \cdot ... \cdot n_k!$ ways to order the element within the chunks.
Hence the multinomial coefficient
\begin{equation*}
    {m \choose n_1, ..., n_k}
\end{equation*}
gives the number of ways to distribute the elements of $[m]$ into ordered buckets of sizes $n_1, ..., n_k$.
A partition corresponds to unordered buckets, i.e. swapping chunks of equal size does not change the partition.
There are $1/\gamma(n_1, ..., n_k)$ ways to do so, and hence we get the expression mentioned above.

@@

Finally, there is still one interesting thought I had while thinking about the above inequality.
Note that the decomposition
\begin{equation*}
    \Pi_m = \bigcup_{k \geq 1, \ n_1 \leq ... \leq n_k} \{ \pi \in \Pi_m \ | \ |\pi| = k, |\{ S \in \pi \ | \ |S| = j \}| = n_j \}
\end{equation*}
corresponds exactly to the core $\mathcal{C}(\Pi_m)$ of $\Pi_m$ as discussed above.
In particular, the decomposition is exactly
\begin{equation*}
    \Pi_m = \bigcup \ \{ \mathrm{Aut}(\Pi_m).\pi \ | \ \pi \in \Pi_m \}
\end{equation*}

>>>>

\title{Miniproject - Combinatorics}
\author{Candidate Number: 1059926}
\date{19.12.2021}

@@

    Consider any $W = \{ (1, i_1, a_1), ..., (1, i_w, a_w) \} \subseteq V_1$ and let $F = \{ a_1, ..., a_w \}$.
    Then
    \begin{equation*}
        W \subseteq \{1\} \times [t] \times F \ \Rightarrow \ |W| \leq t|F|
    \end{equation*}
    By choice of $C_1, ..., C_t$, we have that the number of $j$ with $x \in C_j$ is the same for all $x \in L_{n - 1}$, say $k$.
    Since the $C_i$ are maximal chains, each intersects $L_{n - 1}$ in exactly one element.
    So have bijection
    \begin{align*}
        \{ (x, i) \ | \ x \in L_{n - 1}, 1 \leq i \leq t, x \in C_i\} \to \{ i \ | \ 1 \leq i \leq t \}, \quad (x, i) \mapsto i
    \end{align*}
    where the set on the left-hand side has size $k |L_{n - 1}|$. It follows that $k|L_{n - 1}| = t$.

    Since $\max C_j \in L_{n - 1}$ for all $j$, we have
    \begin{align*}
        N(W) =& \{ (2, j, b) \ | \ b \in L_n, \exists a \in F: \max C_j < a \} \\
        =& \{2\} \times \{ j \ | \ \exists a \in F: \max C_j < a \} \times L_n \\
        =& \{2\} \times \{ j \ | \ \max C_j \in \partial F \} \times L_n
    \end{align*}
    and so by the above
    \begin{align*}
        |N(W)| = |L_n| \sum_{\max C_j \in \partial F} 1 = |L_n| \sum_{x \in \partial F} \sum_{x \in C_j} 1 = |L_n| \sum_{x \in \partial F} k = |L_n| |\partial F| k
    \end{align*}
    Together with the assumption that $|\partial F| |L_n| \geq |F| |L_{n - 1}|$ we see that
    \begin{equation*}
        |W| \leq t |F| = k |L_{n - 1}| |F| \leq k |\partial F| |L_n| = |N(W)|
    \end{equation*}
    So Hall's theorem yields a matching $M \subseteq E$ from $V_1$ to $V_2$.
    As $|V_1| = |V_2|$, this is already a 1-to-1 correspondence.

@@

    Note that this is the partition into \emph{exactly} $k$ sets; 
    I am not completely sure what the exercise description tells us to show, but if it is about the number of ways of decomposing $[m]$ into \emph{at most} $k$ sets, then have
    \begin{equation*}
        \Sigma S(m, k)(2) = 2^{m - 1} \quad \text{and} \quad \Sigma S(m, k)(3) = 3^{m - 1} = \frac 1 2 3^{m - 1} + \frac 1 2
    \end{equation*}
    for $m \geq 3$.
    Here $\Sigma S(m, k) = \sum_{l \leq k} S(m, l)$ is the number of partitions of $[m]$ into at most $k$ sets.
    \begin{proof}
        By Prop.~\ref{prop:partitions_recursion} have that for $m \geq 1$ we have $S(m, 1) = 1$.
        Further, for $m \geq 2$ have
        \begin{equation*}
            S(m, 2) = \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 = 2^{m - 1} - 1
        \end{equation*}
        Applying this once more yields for $m \geq 3$ that
        \begin{align*}
            S(m, 3) =& \sum_{n = 2}^{m - 1} {m - 1 \choose n} (2^{n - 1} - 1) = \frac 1 2 \sum_{n = 1}^{m - 1} {m - 1 \choose n} 2^n - \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 \\
            =&\frac 1 2 \sum_{n = 0}^{m - 1} {m - 1 \choose n} 2^n - \frac 1 2 - \sum_{n = 0}^{m - 1} {m - 1 \choose n} 1 + 1 \\
            =& \frac 1 2 (1 + 2)^{m - 1} - 2^{m - 1} + \frac 1 2 = \frac 1 2 (3^{m - 1} - 2^m + 1)
        \end{align*}
    \end{proof}
    Let $m$ be even and consider $\mathcal{A} \subseteq \Pi_m$ of partitions of $[m]$ into two equally sized sets.
    Then
    \begin{equation*}
        |\mathcal{A}| = \frac 1 2 {m \choose m/2}
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{equation*}
            f: [m]^{(m/2)} \to \mathcal{A}, \quad A \mapsto \{ A, [m] \setminus A \}
        \end{equation*}
        Then for $A$ and $B \neq A, [m] \setminus A$ have that $\{ A, [m] \setminus A \} \neq \{ B, [m] \setminus B \}$.
        Conversely, $A$ and $[m] \setminus A$ have the same image under $f$.
        This shows that $f$ is 2-to-1 and thus 
        \begin{equation*}
            |\mathcal{A}| = \frac 1 2 \left| [m]^{(m/2)} \right| = \frac 1 2 {m \choose m/2}
        \end{equation*}
    \end{proof}
    Further have
    \begin{equation*}
        |\partial \mathcal{A}| = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{align*}
            g: \underbrace{\{ (A, B, C) \ | \ A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m \}}_{ =: \mathcal{G} } \to \partial \mathcal{A}, \quad (A, B, C) &\mapsto \{ A, B, C \}
        \end{align*}
        Note that $g$ is well-defined, as for $A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m$ have that $\pi := \{ A, B, C \}$ is a partition in $L_{m - 3}$ that refines the partition $\{ A, B \cup C \} \in \mathcal{A} \subseteq L_{m - 2}$.
        Further, it is easy to see that $g$ is surjective, as every partition $\pi \in \partial \mathcal{A}$ satisfies $|\pi| = 3$ and has some $A \in \pi$ with $|A| = m/2$.

@@

\section{Part VI}
First of all, in Example~\ref{ex:partitions_have_no_lym} we have already shown that the poset $\Pi_m$ does not satisfy the conditions from Prop.~\ref{prop:characterization_lym}.
Indeed, they do not satisfy (2) of Prop.~\ref{prop:characterization_lym} in a very strong sense:
There is no constant $\epsilon > 0$ such that
\begin{equation*}
    \frac {|\partial F|} {|L_{i - 1}|} \geq \epsilon \frac {|F|} {|L_i|}
\end{equation*}
for all $F \subseteq L_i$.

Now we want to investigate whether $\mathcal{P}_{k, d}$ does. 
\begin{lemma}
    \label{prop:lp_algorithm}
    Let $P$ be a poset with layers $L_0, ..., L_d$.
    Fix $0 < l \leq d$ and let $L_{l - 1} = \{ u_1, ..., u_m \}, L_l = \{ v_1, ..., v_n \}$.

    Assume there exists a matrix $A = (a_{ij}) \in \Q_{\geq 0}^{m \times n}$ such that for all $i, j$
    \footnote{We use $1 = (1, ..., 1)^T$ as the vector consisting of $1$ only here.}
    \begin{equation*}
        e_i^T A 1 = \frac 1 {|L_{l - 1}|} \quad \text{and} \quad 1^T A e_j = \frac 1 {|L_l|}
    \end{equation*}
    and $a_{ij} = 0$ whenever $u_i \not< v_j$ holds.

@@

    Now consider a (finite) sequence of maximal chains $(C_k)_{k \in I}$ such that the chain corresponding to $u_i < v_j$ occurs $a_{ij} |I|$ times (this exists as all the $a_{ij}$ are rational).
    By the assumption, we know that for each $i$ have
    \begin{equation*}
        |\{k \ | \ u_i \in C_k\}| = \sum_{u_i < v_j} a_{ij}|I| = \frac 1 {|L_{l - 1}|} |I|
    \end{equation*}
    and so each element $u_i \in L_{l - 1}$ is contained in the same number $|I| / |L_{l - 1}|$ of chains $C_k$.

    Similarly, for each $J$ have
    \begin{equation*}
        |\{k \ | \ v_j \in C_k\}| = \sum_{v_j > u_i} a_{ij}|I| = \frac 1 {|L_l|} |I|
    \end{equation*}
    The claim follows.
\end{proof}
This lemma does not seem super useful at first, but it gives us a way to check in time polynomial in $O(|P|)$ whether a given poset $P$ satisfies the conditions from Prop.~\ref{prop:characterization_lym}, by using an LP solver.
Note that this is not easily possible when working directly with any of the three conditions defined before.

@@

    If there exists a sequence of maximal chains $C'_1, ..., C'_t$ in $\mathcal{C}(P)$ such that for a chain $C$ chosen uniformly at random from among them, have
    \begin{equation*}
        \Pr[A \in C] = \frac {|A|} {|L_{r(A)}|}
    \end{equation*}
    for all $A \in \mathcal{C}(P)$, then $P$ satisfies the conditions of Prop.~\ref{prop:characterization_lym}.
\end{proposition}
\begin{proof}
    Exactly as in Prop.~\ref{prop:core_of_multisets}.
\end{proof}
I still do not know how to use this to prove that $\mathcal{P}_{k, d}$ does indeed satisfy Prop.~\ref{prop:characterization_lym}, but at least it provides a way for an even faster algorithm to test it for specific $k, d$.
The idea is just that we can adjust the LP algorithm from Lemma~\ref{prop:lp_algorithm} to check the scaled variant, and so we get a polynomial time algorithm in $|\mathcal{C}_{k, d}|$.
In particular, we have the following.
\begin{lemma}
    Let $L_{l - 1} = \{u_1, ..., u_m\}$ and $L_l = \{v_1, ..., v_n\}$ be two consecutive layers of $\mathcal{C}_{k, d}$.
    Assume there exists a matrix $A = (a_{ij}) \in \Q_{\geq 0}^{m \times n}$ such that for all $i, j$
    \begin{equation*}
        e_i^T A 1 = \frac {|P(u_i)|} {\sum_{u \in L_{l - 1}} |P(u)|} \quad \text{and} \quad 1^T A e_j = \frac {|P(v_j)|} {\sum_{v \in L_l} |P(v)|}
    \end{equation*}
    and $a_{ij} = 0$ whenever $u_i \not< v_j$ holds.

@@

    \paragraph{Case 3} If $M = (k_1, ..., k_d)$, see that $2l - 1 = \sum_i k_i$ and so the layers $L_{l - 1}, L_l$ must be the two unique middle layers.
    Now note that $|L_l| = |L_{l - 1}|$ and so it suffices to show that $|\partial F| \geq |F|$.
    However, this follows directly by the symmetric chain decomposition of the poset $P$.
    We have proven the symmetric chain decomposition only for $k_1 = ... = k_d$, but the general version can be found in \cite{anderson}.
\end{proof}
There are two parts I find interesting in this proof.
\begin{itemize}
    \item First of all, the idea to consider all $\bigtimes_i \{ 0, ..., k_i \}$ with possibly distinct $k_i$ makes it easy to use induction, as it suffices to decrease just one coordinate to use the induction hypothesis.
    \item The decomposition
    \begin{equation*}
        L_{i - 1} \times F = \bigcup_{M \in P', r(M) = 2i - 1} \{ (u, a) \in L_{i - 1} \times F \ | \ u + a = M \}
    \end{equation*}
    is just very clever, as it really transforms the set of tuples $(u, a)$ into multiple sets that are more or less $\{ a \in F \ | \ a \leq M \}$ and much easier to handle.
\end{itemize}
Interestingly, I have also found a variant of the LYM inequality (``pseudo-LYM inequality'') for the poset $\Pi_m$ that we studied earlier.
\begin{proposition}
    \label{prop:lym_variant_partitions}
    Let $\mathcal{A} \subseteq \Pi_m$ be a nonempty antichain. Then define
    \begin{equation*}
        a(n_1, ..., n_k) := |\bigl\{ A \in \mathcal{A} \ \bigm| \ |A| = k, \ |\{ S \in A \ | \ |S| = j \}| = |\{ i \ | \ n_i = j \}| \bigr\}|
    \end{equation*}
    the number of partitions $A \in \mathcal{A}$ such that the classes in $A$ have sizes $n_1, ..., n_k$.
    Then
    \begin{equation*}
        \sum_{k \geq 1} \sum_{n_1, ..., n_k} \frac {a(n_1, ..., n_k) k!} {{m \choose n_1, ..., n_k} {m - 1 \choose k - 1}} \leq 1
    \end{equation*}
\end{proposition}
\begin{proof}
    See \cite{lym_partitions}.
\end{proof}
Note that the total number of partitions $\pi$ such that the classes in $\pi$ have sizes $n_1, ..., n_k$ is given by an expression involving the multinomial coefficient, namely
\begin{align*}
    &|\bigl\{ \pi \in \Pi_m \ \bigm| \ |\pi| = k, |\{ S \in \pi \ | \ |S| = j\}| \\
    = &|\{ i \ | \ n_i = j \}| \bigr\}| = {m \choose n_1, ..., n_k} \underbrace{ \prod_i \frac 1 {|\{ j \ | \ n_j = i\}|!} }_{=: \gamma(n_1, ..., n_k)}
\end{align*}
Intuitively, the reason is that there are $m!$ ways to order the elements $[m]$, and after splitting such a sequence in chunks of sizes $n_1, ..., n_k$, there are $n_1! \cdot ... \cdot n_k!$ ways to order the element within the chunks.
Hence the multinomial coefficient
\begin{equation*}
    {m \choose n_1, ..., n_k}
\end{equation*}
gives the number of ways to distribute the elements of $[m]$ into ordered buckets of sizes $n_1, ..., n_k$.
A partition corresponds to unordered buckets, i.e. swapping chunks of equal size does not change the partition.
There are $1/\gamma(n_1, ..., n_k)$ ways to do so, and hence we get the expression mentioned above.

@@

Finally, there is still one interesting thought I had while thinking about the above inequality.
Note that the decomposition
\begin{equation*}
    \Pi_m = \bigcup_{k \geq 1, \ n_1 \leq ... \leq n_k} \bigl\{ \pi \in \Pi_m \ | \ |\pi| = k, |\{ S \in \pi \ | \ |S| = j \}| = |\{ i \ | \ n_i = j \}| \bigr\}
\end{equation*}
corresponds exactly to the core $\mathcal{C}(\Pi_m)$ of $\Pi_m$ as discussed above.
In particular, the decomposition is exactly
\begin{equation*}
    \Pi_m = \bigcup \ \{ \mathrm{Aut}(\Pi_m).\pi \ | \ \pi \in \Pi_m \}
\end{equation*}
